Tomographic galaxy clustering with the Subaru Hyper Suprime-Cam first year public data release

### 2019-10-23 doc/Paper/main.tex
affiliations and acknowledgements

\author[f]{Humna Awan,}
\author[f]{Adam Broussard,}
\author[b]{Zahra Gomes,}
\author[f]{Eric Gawiser,}
\author[x]{Rachel Mandelbaum,}
\author[g,h,i,j]{Hironao Miyatake,}
\author[k]{Jeffrey A. Newman,}
\author[x]{Ignacio Sevilla,}
\author[e,l]{Sarah Skinner,}
\author[x]{Erica Wagoner}

\affiliation[a]{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton NJ 08544-0010, USA}
\affiliation[b]{Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, United Kingdom}
\affiliation[c]{Fermi National Accelerator Laboratory, Batavia, IL, 60510, USA}
\affiliation[d]{Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA}
\affiliation[e]{Brookhaven National Laboratory, Physics Department, Upton, NY 11973, USA}
\affiliation[f]{Rutgers University, Physics \& Astronomy Department, Piscataway, NJ 08854, USA}
\affiliation[g]{Institute for Advanced Research, Nagoya University, Nagoya 464-8601, Japan}
\affiliation[h]{Division of Physics and Astrophysical Science, Graduate School of Science, Nagoya University, Nagoya 464-8602, Japan}
\affiliation[i]{Kavli Institute for the Physics and Mathematics of the Universe (Kavli IPMU, WPI), UTIAS, The University of Tokyo, Chiba 277- 8583, Japan}
\affiliation[j]{Jet Propulsion Laboratory, California Institute of Technology, Pasadena, CA 91109, USA}
\affiliation[k]{Department of Physics and Astronomy and PITT PACC, University of Pittsburgh, Pittsburgh, PA, 15260, USA}
\affiliation[l]{Department of Physics, Missouri University of Science and Technology, Rolla, MO 65401, USA}

@@

\section{Introduction}\label{sec:intro}

The past two decades have seen a revolution in our understanding of the Universe and its constituents. While observations of the cosmic microwave background (CMB) play a pivotal role in anchoring the standard  cosmological model \cite{Planck:2018}, they are not a direct probe of the physics of the low redshift Universe. In order to directly characterize properties of dark energy and modified gravity, we need to measure the expansion history and growth  \cite{1903.12016}. At the moment, this is only possible with optical large scale structure surveys.

@@

The community is well aware of these challenges lying ahead. There are two main approaches to preparing for the arrival of data. On one hand, we are building sophisticated, realistic mock data sets, such as LSST DESC Data Challenge 2 \cite{1909.07340,1907.06530}. On the other hand, we are (re-)analyzing existing precursor data to test-drive our methods and codes in a realistic environment. In this paper, we focus on the latter approach and perform an analysis of photometric galaxy clustering using the first data release (DR1) of the Hyper Suprime-Cam Subaru Strategic Program (HSC-SSP) \cite{2018PASJ...70S...8A}. This dataset bears significant similarities to the data expected from LSST, both in terms of survey depth and photometric bands as well as primary data reduction and catalog generation. Specifically, HSC covers essentially the same bands as LSST, with the exception of missing the UV $u$-band fluxes. In this work, we focus on a magnitude-limited galaxy sample with limiting magnitude $i<24.5$, which is similar to the expected 5-$\sigma$ detection limit for LSST after one year \cite{1809.01669} ($i<25.1$). Finally, the primary data reduction codes used in HSC-SSP and planned for LSST are of the same lineage, employing much of the same methods \cite{2018PASJ...70S...5B,1812.03248}. With the exception of a rather small sky area of around 90 square degrees used in this paper, the HSC-SSP data is therefore a perfect proxy for future LSST data. This analysis has the additional attraction that no other photometric galaxy clustering study has been carried out with these data.

@@

  Despite the small fraction of sky covered by the data used in this work, HSC DR1 constitutes an ideal proxy for the data expected for LSST. Our analysis demonstrates the feasibility of photometric galaxy clustering analyses using such high signal-to-noise datasets, which bodes well for future cosmological constraints using data from LSST.

\acknowledgments

DA acknowledges support from the Beecroft trust and from STFC through
an Ernest Rutherford Fellowship, grant reference ST/P004474/1.  HA, BA
and EG have been supported by the Rutgers Discovery Informatics
Institute Fellowship of Excellence in Computational and Data Science
(AY 2017-2020) and Rutgers University \& Bevier Dissertation
Completion Fellowship (AY 2019-2020). HA, AB, and EG were supported by
the Department of Energy (grants DE-SC0011636 and DE-SC0010008). HA
also thanks the LSSTC Data Science Fellowship Program, which is funded
by LSSTC, NSF Cybertraining Grant \#1829740, the Brinson Foundation,
and the Moore Foundation, as participation in the program has
benefited this work. ZG is supported by a Rhodes Scholarship granted
by the Rhodes Trust. HM has been supported by the Jet Propulsion Laboratory,
California Institute of Technology, under a contract with the National
Aeronautics and Space Administration and Grant-in-Aid for Scientific
Research from the JSPS Promotion of Science (No. 18H04350). 
JAN us supported by grant DOE DE-SC0007914 from
the U.S. Department of Energy Office of Science (Office of High Energy
Physics). This manuscript has been authored by Fermi Research
Alliance, LLC under Contract No. DE-AC02-07CH11359 with the
U.S. Department of Energy, Office of Science, Office of High Energy
Physics. The work by JS was supported by the U.S. Department of Energy
award DE-SC0009920. This project was supported in part by the
U.S. Department of Energy, Office of Science, Office of Workforce
Development for Teachers and Scientists (WDTS) under the Science
Undergraduate Laboratory Internships Program (SULI).

@@

This paper makes use of software developed for the Large Synoptic Survey Telescope. We thank the LSST Project for making their code available as free software at  http://dm.lsst.org

>>>>

\author[f]{Humna Awan,}
\author[f]{Adam Broussard,}
\author[b]{Zahra Gomes,}
\author[f]{Eric Gawiser,}
\author[g]{Rachel Mandelbaum,}
\author[h,i,j,k]{Hironao Miyatake,}
\author[l]{Jeffrey A. Newman,}
\author[m]{Ignacio Sevilla,}
\author[e,n]{Sarah Skinner,}
\author[o]{Erica Wagoner}

\affiliation[a]{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton NJ 08544-0010, USA}
\affiliation[b]{Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, United Kingdom}
\affiliation[c]{Fermi National Accelerator Laboratory, Batavia, IL, 60510, USA}
\affiliation[d]{Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA}
\affiliation[e]{Brookhaven National Laboratory, Physics Department, Upton, NY 11973, USA}
\affiliation[f]{Rutgers University, Physics \& Astronomy Department, Piscataway, NJ 08854, USA}
\affiliation[g]{McWilliams Center for Cosmology, Department of Physics, Carnegie Mellon University, Pittsburgh, PA 15213, USA}
\affiliation[h]{Institute for Advanced Research, Nagoya University, Nagoya 464-8601, Japan}
\affiliation[i]{Division of Physics and Astrophysical Science, Graduate School of Science, Nagoya University, Nagoya 464-8602, Japan}
\affiliation[j]{Kavli Institute for the Physics and Mathematics of the Universe (Kavli IPMU, WPI), UTIAS, The University of Tokyo, Chiba 277- 8583, Japan}
\affiliation[k]{Jet Propulsion Laboratory, California Institute of Technology, Pasadena, CA 91109, USA}
\affiliation[l]{Department of Physics and Astronomy and PITT PACC, University of Pittsburgh, Pittsburgh, PA, 15260, USA}
\affiliation[m]{Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
\affiliation[n]{Department of Physics, Missouri University of Science and Technology, Rolla, MO 65401, USA}
\affiliation[o]{Department of Physics, University of Arizona, Tucson, AZ, 85721, USA}

@@

\section{Introduction}\label{sec:intro}
  The past two decades have seen a revolution in our understanding of the Universe and its constituents. While observations of the cosmic microwave background (CMB) play a pivotal role in anchoring the standard  cosmological model \cite{Planck:2018}, they are not a direct probe of the physics of the low redshift Universe. In order to directly characterize properties of dark energy and modified gravity, we need to measure the expansion history and growth  \cite{1903.12016}. At the moment, this is only possible with optical large scale structure surveys.

@@

  The community is well aware of these challenges lying ahead. There are two main approaches to preparing for the arrival of data. On one hand, we are building sophisticated, realistic mock data sets, such as LSST Dark Energy Science Collaboration (DESC) Data Challenge 2 \cite{1909.07340,1907.06530}. On the other hand, we are (re-)analyzing existing precursor data to test-drive our methods and codes in a realistic environment. In this paper, we focus on the latter approach and perform an analysis of photometric galaxy clustering using the first data release (DR1) of the Hyper Suprime-Cam Subaru Strategic Program (HSC-SSP) \cite{2018PASJ...70S...8A}. This dataset bears significant similarities to the data expected from LSST, both in terms of survey depth and photometric bands as well as primary data reduction and catalog generation. Specifically, HSC covers essentially the same bands as LSST, with the exception of missing the UV $u$-band fluxes. In this work, we focus on a magnitude-limited galaxy sample with limiting magnitude $i<24.5$, which is similar to the expected 5-$\sigma$ detection limit for LSST after one year \cite{1809.01669} ($i<25.1$). Finally, the primary data reduction codes used in HSC-SSP and planned for LSST are of the same lineage, employing much of the same methods \cite{2018PASJ...70S...5B,1812.03248}. With the exception of a rather small sky area of around 90 square degrees used in this paper, the HSC-SSP data is therefore a perfect proxy for future LSST data. This analysis has the additional attraction that no other photometric galaxy clustering study has been carried out with these data.

@@

  Despite the small fraction of sky covered by the data used in this work, HSC DR1 constitutes an ideal proxy for the data expected for LSST. Our analysis demonstrates the feasibility of photometric galaxy clustering studies using such high signal-to-noise datasets. In particular, this work shows the level of readiness of the LSST DESC analysis pipelines, and showcases the application of power spectrum methods for galaxy clustering, and the use of template deprojection techniques to handle contamination from sky systematics. These results therefore bode well for future cosmological constraints using galaxy clustering data from LSST.

\acknowledgments
  DA acknowledges support from the Beecroft trust and from the Science and Technology Facilities Council through an Ernest Rutherford Fellowship, grant reference ST/P004474/1.
  HA, BA and EG have been supported by the Rutgers Discovery Informatics Institute Fellowship of Excellence in Computational and Data Science (AY 2017-2020) and Rutgers University \& Bevier Dissertation Completion Fellowship (AY 2019-2020).
  HA, AB, and EG were supported by the Department of Energy (grants DE-SC0011636 and DE-SC0010008).
  HA also thanks the LSSTC Data Science Fellowship Program, which is funded by LSSTC, NSF Cybertraining Grant \#1829740, the Brinson Foundation, and the Moore Foundation, as participation in the program has benefited this work.
  ZG is supported by a Rhodes Scholarship granted by the Rhodes Trust.
  RM is supported by the Department of Energy Cosmic Frontier program, grant DE-SC0010118.
  HM has been supported by the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the National Aeronautics and Space Administration and Grant-in-Aid for Scientific Research from the JSPS Promotion of Science (No. 18H04350). 
  JAN us supported by grant DOE DE-SC0007914 from the U.S. Department of Energy Office of Science (Office of High Energy Physics).
  This manuscript has been authored by Fermi Research Alliance, LLC under Contract No. DE-AC02-07CH11359 with the U.S. Department of Energy, Office of Science, Office of High Energy Physics.
  The work by JS was supported by the U.S. Department of Energy award DE-SC0009920.
  This work was supported by the grant MDM-2015-0509 from the Spanish Ministry of Science and Innovation.
  This project was supported in part by the U.S. Department of Energy, Office of Science, Office of Workforce Development for Teachers and Scientists (WDTS) under the Science Undergraduate Laboratory Internships Program (SULI).

@@

  This paper makes use of software developed for the Large Synoptic Survey Telescope. We thank the LSST Project for making their code available as free software at \url{http://dm.lsst.org}.
